---
title: "Population structure and insecticide response of Gammarus spp. in agricultural and upstream forested sites of small streams "
author: "Schneeweiss A., Schreiner V.C., Liess M., Roeder N., Schwenk K., Schaefer R.B."
date: "01 November 2022"
output:
  html_document:
    figure_caption: yes
    highlight: tango
    number_sections: no 
    toc: yes
    toc_float: 
         toc_collapsed: true
    toc_depth: 4
  pdf_document:
    toc: yes
    toc_float: 
         toc_collapsed: true
    toc_depth: '4'
  word_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

R script written by Anke Schneeweiss

Internally verified by Dr. Verena C. Schreiner

Institute for Environmental Sciences
RPTU Kaiserslautern-Landau 
Fortstrasse 7  
76829 Landau  
GERMANY  
Corresponding mail address: schneeweiss'@'uni-landau.de

```{r, include=FALSE}
#########Load packages and library files##################
library(lattice)  
library(lme4)
require(dplyr)
require(tidyr)
#source(file = ".../HighstatLibV10.R") #download file "GLMGLMM_RCode.zip" on https://www.highstat.com/index.php/books2?view=article&id=21&catid=18 and read in "HighstatLibV10.R" from local source
library(numDeriv)
library(ggplot2)
library(TMB)
library(glmmTMB)
library(bbmle)
library(data.table)
library(ggpubr)
library(optimx) 
library(effects)
library(ggeffects)
library(see) 
library(emmeans) 
library(multcomp)#
library(car)
library(sjPlot)
library(snakecase)
library(knitr)
library(kableExtra) 
library("performance")
library(qqplotr)
library(DHARMa)
require(beanplot)
require(patchwork)
require(jtools)
library(ggbreak)
library(grid)
```

```{r, include=FALSE}
#########Set the working directory ##########
path<-getwd()
#Set data input path
Inpath<-paste(path,"0_data_input" ,sep="/")
#set analysis output path
Outpath<-paste(path,"1_analysis_output" ,sep="/")
```

# 0. Some general definitions
We define a fixed color scheme for the site types.
```{r}
#group color scheme for site_types in all plots
group.color <- c("agriculture"= "peru", "edge" ="darkkhaki" , "refuge" = "darkgreen")
```
Round 0.5 up to 1.
```{r}
round2 <- function(x, digits = 0) {  
  posneg <- sign(x)
  z <- abs(x) * 10^digits
  z <- z + 0.5
  z <- trunc(z)
  z <- z / 10^digits
  z * posneg
}
```

# 1. Hypothesis 1: Pesticide tolerance.
Given the regular pesticide exposure at agricultural sections in our study region, we hypothesised local adaptation processes in terms of higher pesticide tolerance of Gammarus spp. and that these propagate via gene flow to the edge habitats (hypothesis 1). 

## 1.1 Load and prepare acute toxicity test data

```{r}
df1 <- read.csv(file=paste(Inpath,"20230505_data_part_1_tolerance.csv",sep="/"),header=TRUE, sep=";", na.strings =c("na","NA",""),stringsAsFactors = FALSE)
df2<-read.csv(file=paste(Inpath,"20230505_data_4_gammarus_abundance.csv",sep="/"),header=TRUE, sep=";", na.strings ="na",stringsAsFactors = FALSE)
```
df1: data of acute toxicity test results for 18 sites (mortality and immobility of Gammarus spp. after 96h imidacloprid exposure)
df2: data of relative Gammarus species abundances for the 18 sites


Column definitions for df1:
Column 1-3: "Site.id", "Stream" and "Site type" - sampling site characterization
Column 4: "Date" Date of Gammarid sampling and test start
Column 5: "Concentration" Nominal imidacloprid concentration used in the test in µg/L; 7-8 test concentrations (control included)
Column 6: "Treatment" Treatment A-G,Z according to concentrations
Column 7: "Replicate" Test vessel replicate, usually 4 per concentration and sampling site
Column 8: "Individuals" The number of individuals per vessel at the start of the test
Column 9: "Alive_96" Number of living individuals monitored 96h after test start (immobile organism are included)
Column 11: "Immobile_96" Number of immobile individuals monitored 96h after test start; immobile refers to slight movements of antenna or legs allowed, but no swimming --> dead organisms are not included in the recorded response immobility --> but dead+immobile are summed up for model immobility
Column 11: "Dead_96" Number of dead individuals monitored 96h after test start
Column 12: "abundance" describes the relative abundance of Gammarus fossarum at the respective sampling site (high abundance: >=80% Gammarus fossarum; low: <=50% Gammarus fossarum)

```{r}
#########Prepare data############################
df3<-df1
df3$Site=tolower(df3$Site)
str(df3)

#Response variable is binary: dead or alive, immobile or mobile 
df3$Proportion_dead_96<-df3$Dead_96/df3$Individuals # calculate mortality proportion

#Explanatory variables
#fixed factors
df3$fSiteType  	<- factor(df3$Site) # in manuscript referred to as "site type", categorical, 3 levels - agriculture-edge-refuge
df3$LogConcentration <- log10(df3$Concentration+3.5) # continuous, log-transformed as it has a large scale (+0.5*lowest non-zero concentration, to avoid infinite numbers)
#random factors
df3$fStream 		<- factor(df3$Stream) # categorical, 6 levels
df3$fGFossarumAbundance	<- factor(df3$abundance)# categorical, 2 levels
df3 <- unite(df3,"replicate_subunit",c("fStream","fSiteType","Treatment"),sep="",remove = F)#prepare unique identifier for true replicates to avoid pseudo-replication
df3$fSubunit		<- factor(df3$replicate_subunit)
```

```{r}
##Subsetting for different models
df_gam<-df3

#without outlier Hainbach_agriculture_Control_0_ug/L-all dead; since it does not change results, we keep the outlier
#df_gam_outl<-df_gam[-c(15),]
#df_gam<-df_gam %>% subset(!Stream=="Otterbach"&Treatment=="E")
```

## 1.2 Gammarus spp. 
### 1.2.1 Mortality after 96h imidacloprid exposure - Predictor Site type
#### 1.2.1.1 Statistics
```{r, echo=FALSE}
#statistics for the following dataframe
df_statistics<-df_gam %>% dplyr::select(c("Alive_96","Dead_96","Proportion_dead_96","Individuals","LogConcentration","fSiteType","fStream","fGFossarumAbundance","fSubunit"))
str(df_statistics)
# remove NAs
df_statistics<-df_statistics[!is.na(df_statistics$Proportion_dead_96),]
unique(df_statistics$Proportion_dead_96)
```

a) Data exploration
```{r}
#outliers
MyVar <- c("Alive_96","Dead_96","Proportion_dead_96","LogConcentration")
Mydotplot(df_statistics[,MyVar]) 
outlier_values <- boxplot.stats(df_statistics$Proportion_dead_96)$out  # outlier values.
boxplot(df_statistics$Proportion_dead_96, main="outliers", boxwex=0.1)
mtext(paste("Outliers: ", paste(outlier_values, collapse=", ")), cex=0.6)
#variance homogeneity
p_vh<-beanplot(Proportion_dead_96~fSiteType, data=df_statistics)
leveneTest(Proportion_dead_96~fSiteType, data=df_statistics)
#normal distribution
hist(df_statistics$Proportion_dead_96)
shapiro.test(df_statistics$Proportion_dead_96)
#Inspect for collinearity with multi-panel scatterplots (only covariates): 
MyVar1 <- c("LogConcentration","fGFossarumAbundance","fSiteType") # multi-panel scatterplots for fixed factors
pairs(df_statistics[,MyVar1], 
    upper.panel = panel.cor,
  diag.panel  = panel.hist,
  lower.panel = panel.smooth)
#Inspect for collinearity with VIF, cut-off level in Zuur et al., 2013 VIF = 3
corvif(df_statistics[ ,c("LogConcentration","fGFossarumAbundance","fSiteType")]) # calculate variance inflation factors(VIFs) for fixed covariates 
```

Binary data.
Note that in the control of Hainbach-agriculture all 5 organism were dead --> rerun whole analysis without this observation does not change the results/conclusion, we keep it. 

The random factor "fGFossarumAbundance" is correlated with the fixed factor "fSiteType" (correlation coefficient = -0.7). Since we are explicitly interested in the fixed covariate fSiteType, we decided to keep fSiteType in the model and to remove fGFossarumAbundance. This must be kept in mind for the interpretation of the model output. To illustrate our explanations within the discussion section of the manuscript, we constructed a second model with fGFossarumAbundance instead of fSiteType as fixed factor. 

b) Modeling
```{r echo=FALSE}
#Executing the GLMM using the glmer function
Full_model <- glmer(Proportion_dead_96 ~ LogConcentration*fSiteType+(1|fStream)+(1|fSubunit),weights = Individuals,
            family = binomial, data = df_statistics,
            glmerControl(optimizer = "optimx", calc.derivs = FALSE,
                         optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)))
```

b) Model selection
Model selection is applied to the fixed structure only, since the random effects are part of the experimental design.
Find optimal fixed structure:
```{r}
drop1(Full_model, test = 'Chisq')#Interaction is not significant, we remove it from the model.

#same without interaction
Reduced_model <- glmer(Proportion_dead_96 ~ LogConcentration+fSiteType+(1|fStream)+(1|fSubunit),weights = Individuals,
            family = binomial, data = df_statistics,
            glmerControl(optimizer = "optimx", calc.derivs = FALSE,
                         optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)))

drop1(Reduced_model, test = 'Chisq')#fSiteType is significant, we keep it 
```

c) Model validation
```{r}
#model validation with the performance package
check_model(Reduced_model) 
model_performance(Reduced_model) 
check_overdispersion(Reduced_model) 
check_singularity(Reduced_model) 
check_collinearity(Reduced_model) 
```
Model is valid.

e) Model output
```{r}
summary(Reduced_model)
# create a nice output
tab_model(Reduced_model,show.ci = F, show.se = TRUE)#,  string.se = "SE") 

#significance of fixed effects
Anova(Reduced_model,type=2)#concentration and site_type are significant, no predictor can be droped
drop1(Reduced_model)
#posthoc - more conservative --> used for main manuscript
emmeans(Reduced_model, list(pairwise ~ fSiteType), adjust = "tukey")
#store in unique name
Model1<-Reduced_model
```

#### 1.2.1.2 LC50
```{r}
#to calculate LC50--> mortality prob = 0.5 resolve model x1 = (logit(prob) - b0-b2x2-b3x3)/b1
#logit(prob) = ln(prob/(1-prob))
#Agriculture: x2=0 and x3=0
#Edge: x2=1 and x3=0
#Refuge: x2=0 and x3=1
# note that log() is the same as ln(), natural logarithm, for base ten, please specify log10()
function_calculate_LC50 = function(b0,b1,b2,b3,prob){
x1 = (-b0-b2-b3+log(-prob/(prob-1)))/b1
#x1 = (log(prob/(1-prob))-b0-b2-b3)/b1 #both lines are the same
return(10^(x1)) # inverse of log10() for the concentration, given that it was log10() transformed
}

#access the estimated parameters and calculate predictor function
ff<-fixef(Reduced_model)
LC_all<-function_calculate_LC50(ff[[1]],ff[[2]],ff[[3]],ff[[4]],.5)
LC_agriculture<-function_calculate_LC50(ff[[1]],ff[[2]],0,0,.5) #LC50 agriculture
LC_edge<-function_calculate_LC50(ff[[1]],ff[[2]],ff[[3]],0,.5) #LC50 edge
LC_refuge<-function_calculate_LC50(ff[[1]],ff[[2]],0,ff[[4]],.5) #LC50 refuge
LC_agriculture # 224 µg/L
LC_edge # 439 µg/L
LC_refuge # 492 µg/L
#x1 = concentration, b0 parameter of agriculture--> intercept, b1 parameter of Lconc, b2 edge, b3 refuge
```

#### 1.2.1.3 Plot
Note: this plot does not take the uncertainty of the random effects parameters into account.
```{r}
#Predict new data (following Zuur et al., 2013)
#create grid of covariate values and get dummy variables
NewData <- expand.grid(fSiteType=c("agriculture","edge","refuge"),
                        LogConcentration=seq(log10(2.3),3.4,length=1000))
#figure out the model matrix X corresponding to the new data
X <- model.matrix(~ LogConcentration + fSiteType, data = NewData)

#access the estimated parameters and calculate predictor function
ff<-fixef(Reduced_model)
NewData$eta         <- X %*% ff

#calculate the probability of mortality for the specified covariate values using the inverse of the logistic link function
NewData$SuccessPred <- exp(NewData$eta) / (1 + exp(NewData$eta)) 

#extract the variance-covariance matrix of the parameter V , compute XVX' to get the variance-covariance matrix of the predictions, then extract the diagonal of this matrix to get variances of predictions
output<-vcov(Reduced_model) #use only for glmmTMB models,called "variance"
NewData$VarPred <- diag(X %*% output %*% t(X))#unconditional variance
#take the square-root to get standard errors of the predictions
NewData$sePred  <- sqrt(NewData$VarPred)

#obtain 95% confidence intervals on the scale of the response variable
#for GLMMs, run the confidence interval boundaries (not the standard errors) through the inverse-link function
NewData$seLow  <- exp(NewData$eta - 1.96 * NewData$sePred) / (1 + exp(NewData$eta - 1.96 * NewData$sePred))
NewData$seHigh <- exp(NewData$eta + 1.96 * NewData$sePred) / (1 + exp(NewData$eta + 1.96 * NewData$sePred))

#aggregate data as mean of all subunit replicates
df_mean<-aggregate(list(df_statistics$Proportion_dead_96, df_statistics$Individuals), by=list(df_statistics$fSubunit, df_statistics$fSiteType,df_statistics$fStream,df_statistics$LogConcentration,df_statistics$fGFossarumAbundance), mean)
str(df_mean)
names(df_mean)<-c("fSubunit","fSiteType","fStream","LogConcentration","fGFossarumAbundance","Proportion_dead_96","Individuals")

#use this one if you want to show the mean of all subunit replicates, to avoid overplotting
df_statistics2  <-df_mean %>% mutate(result=NA, result=ifelse(fSiteType =="agriculture",LogConcentration-0.04,result), result=ifelse(fSiteType =="refuge",LogConcentration+0.04,result), result=ifelse(fSiteType =="edge",LogConcentration,result)  )

p_G_sp_mortality <- 
  ggplot(NewData, mapping=aes(x=LogConcentration, y=SuccessPred,group=fSiteType)) +
  geom_point(df_statistics2,mapping = aes(x=result,y=Proportion_dead_96,colour=fSiteType, alpha = fSiteType),size=1.5)+ # position = position_dodge2(preserve = "single")  
  geom_line(aes(colour=fSiteType),size=1.1, alpha=0.8) +
    geom_ribbon(aes( ymin= seLow, ymax=seHigh, fill = fSiteType, alpha = fSiteType))+
       scale_x_continuous(breaks = c(log10(1),log10(10),log10(100),log10(1000),log10(5000)), labels=c(1,10,100,1000,5000),limits = c(log10(1),log10(5000)),)+
  scale_y_continuous(breaks = c(0,0.20,0.40,0.6,0.80,1), labels=c("0%","20%","40%","60%","80%","100%"),limits = c(0,1))+
  scale_colour_manual(values=group.color,name="Site type",breaks = c("refuge","edge","agriculture"), labels=c("Refuge","Edge","Agriculture"))+
    scale_fill_manual(values=group.color,name="Site type",breaks = c("refuge","edge","agriculture"), labels=c("Refuge","Edge","Agriculture"))+ 
   scale_alpha_manual(values=c(0.2,0.4,0.3),name="Site type",breaks = c("refuge","edge","agriculture"), labels=c("Refuge","Edge","Agriculture"))+
  annotation_logticks(sides = "b",colour = "grey50")+
    labs(x = expression(paste("Imidacloprid concentration [µg/L]")), y = 'Mortality')+
guides(fill = guide_legend(title.position = "top",title.hjust = 0.5),colour= guide_legend(title.position = "top",title.hjust = 0.5))+
   theme_light()+
  theme(  axis.title = element_text(colour = "grey30"),
          axis.text = element_text(colour = "grey50"),
           #axis.ticks.x = element_line(colour = "grey90"),
          panel.border = element_blank(),
          axis.line = element_line(colour = "grey80"),
          panel.grid.major = element_line(colour = "grey90"),
          legend.text = element_text(colour = "grey50"),
          legend.title = element_text(colour = "grey30"),
          legend.position="top")
p_G_sp_mortality
```

### 1.2.2 Mortality after 96h imidacloprid exposure - Predictor G.fossarum Abundance
#### 1.2.2.1 Statistics
```{r, echo=FALSE}
#statistics for the following dataframe
df_statistics<-df_gam %>% dplyr::select(c("Proportion_dead_96","Individuals","LogConcentration","fSiteType","fStream","fGFossarumAbundance","fSubunit"))
str(df_statistics)
unique(df_statistics[,c("fStream","fSiteType","fGFossarumAbundance")])
```

a) Modelling
```{r echo=FALSE}
#Executing the GLMM using the glmer function
Full_model <- glmer(Proportion_dead_96 ~ LogConcentration*fGFossarumAbundance+(1|fStream)+(1|fSubunit),weights = Individuals,
            family = binomial, data = df_statistics,
            glmerControl(optimizer = "optimx", calc.derivs = FALSE,
                         optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)))
```

b) Model selection
Find optimal fixed structure:
```{r}
Anova(Full_model)
drop1(Full_model)#Interaction is not significant, we remove it from the model.

#same without interaction
Reduced_model <- glmer(Proportion_dead_96 ~ LogConcentration+fGFossarumAbundance+(1|fStream)+(1|fSubunit),weights = Individuals,
            family = binomial, data = df_statistics,
            glmerControl(optimizer = "optimx", calc.derivs = FALSE,
                         optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)))
Anova(Reduced_model)
drop1(Reduced_model)#fGFossarumAbundance is significant, we keep it
```

c) Model validation
```{r}
#model validation with the performance package
check_model(Reduced_model) 
model_performance(Reduced_model) 
check_overdispersion(Reduced_model) 
check_singularity(Reduced_model) 
check_collinearity(Reduced_model) 
icc(Reduced_model)
```
Model is valid.

e) Model output
```{r}
summary(Reduced_model)
#output
tab_model(Reduced_model,show.ci = F, show.se = TRUE)#,  string.se = "SE")

#significance of fixed effects
Anova(Reduced_model,type=2)#concentration and site_type are significant, no predictor can be droped
drop1(Reduced_model)
#posthoc - more conservative --> used for main manuscript
emmeans(Reduced_model, list(pairwise ~ fGFossarumAbundance), adjust = "tukey")
#store in unique name
Model2<-Reduced_model
```
Results: The difference observed between site types may be explained by the differences in the relative abundance of Gammarus fossarum at the individual site types.

```{r}
#calculate LC50
function_calculate_LC50 = function(b0,b1,b2,prob){
x1 = (-b0-b2+log(-prob/(prob-1)))/b1
#x1 = (log(prob/(1-prob))-b0-b2-b3)/b1 #both lines are the same
return(10^(x1)) # inverse of log10() for the concentration, given that it was log10(x+1) transformed
}

#access the estimated parameters and calculate predictor function
ff<-fixef(Reduced_model)
LC_high<-function_calculate_LC50(ff[[1]],ff[[2]],0,.5) #LC50 high abundance
LC_low<-function_calculate_LC50(ff[[1]],ff[[2]],ff[[3]],.5) #LC50 low abundance
LC_high # 443 µg/L
LC_low # 189 µg/L
```

#### 1.2.2.3 Plot
```{r}
#Predict new data
#create grid of covariate values and get dummy variables
NewData <- expand.grid(fGFossarumAbundance=c("high","low"),
                        LogConcentration=seq(log10(2.3),3.4,length=1000))
#figure out the model matrix X corresponding to the new data
X <- model.matrix(~ LogConcentration + fGFossarumAbundance, data = NewData)

#access the estimated parameters and calculate predictor function
ff<-fixef(Reduced_model)
NewData$eta         <- X %*% ff

#calculate the probability of mortality for the specified covariate values using the inverse of the logistic link function
NewData$SuccessPred <- exp(NewData$eta) / (1 + exp(NewData$eta)) # same as NewData$SuccessPred <-1 / (1 + exp(-NewData$eta)) ---> inverse: transforms a real number (usually the logarithm of the odds) to a value (usually probability p) in the interval [0,1]. The invlogit function is 1 / (1 + exp(-x)).

#extract the variance-covariance matrix of the parameter V , compute XVX' to get the variance-covariance matrix of the predictions, then extract the diagonal of this matrix to get variances of predictions
output<-vcov(Reduced_model) #use only for glmmTMB models,called "variance"
NewData$VarPred <- diag(X %*% output %*% t(X))#unconditional variance
#take the square-root to get standard deviations (errors) of the predictions
NewData$sePred  <- sqrt(NewData$VarPred)

#obtain 95% confidence intervals on the scale of the response variable
#for GLMMs, run the confidence interval boundaries (not the standard errors) through the inverse-link function
NewData$seLow  <- exp(NewData$eta - 1.96 * NewData$sePred) / (1 + exp(NewData$eta - 1.96 * NewData$sePred))
NewData$seHigh <- exp(NewData$eta + 1.96 * NewData$sePred) / (1 + exp(NewData$eta + 1.96 * NewData$sePred))

#aggregate data as mean of all subunit replicates
df_mean<-aggregate(list(df_statistics$Proportion_dead_96, df_statistics$Individuals), by=list(df_statistics$fSubunit, df_statistics$fSiteType,df_statistics$fStream,df_statistics$LogConcentration,df_statistics$fGFossarumAbundance), mean)
str(df_mean)
names(df_mean)<-c("fSubunit","fSiteType","fStream","LogConcentration","fGFossarumAbundance","Proportion_dead_96","Individuals")

#use this one if you want to show the mean of all subunit replicates, to avoid overplotting
df_statistics2  <-df_mean %>% mutate(result=NA, result=ifelse(fGFossarumAbundance =="low",LogConcentration,result), result=ifelse(fGFossarumAbundance  =="high",LogConcentration+0.02,result))

p_G_sp_mortality_agric <- 
  ggplot(NewData, mapping=aes(x=LogConcentration, y=SuccessPred,group=fGFossarumAbundance)) +
  geom_point(df_statistics2,mapping = aes(x=result,y=Proportion_dead_96,colour=fGFossarumAbundance),size=1.5, alpha = 0.3)+ # position = position_dodge2(preserve = "single")  
  geom_line(aes(colour=fGFossarumAbundance),size=1.2, alpha=1) +
    geom_ribbon(aes( ymin= seLow, ymax=seHigh, fill = fGFossarumAbundance), alpha = 0.2 )+
     scale_x_continuous(breaks = c(log10(1),log10(10),log10(100),log10(1000),log10(5000)), labels=c(1,10,100,1000,5000),limits = c(log10(1),log10(5000)))+
  scale_y_continuous(breaks = c(0,0.20,0.40,0.6,0.80,1), labels=c("0%","20%","40%","60%","80%","100%"),limits = c(0,1))+
 scale_fill_brewer(type = "seq", palette="Set1",name=expression(paste("Abundance of" ,phantom(x), italic("Gammarus  fossarum"))),breaks = c("low","high"), labels=c("Low (≤50%, n=4)", "High (≥80%, n=14)"))+
  scale_colour_brewer(type = "seq", palette="Set1",name=expression(paste("Abundance of" ,phantom(x), italic("Gammarus  fossarum"))),breaks = c("low","high"), labels=c("Low (≤50%, n=4)", "High (≥80%, n=14)"))+
  annotation_logticks(sides = "b",colour = "grey50")+
    labs(x = expression(paste("Imidacloprid concentration [µg/L]")), y = 'Mortality')+
    guides(fill=guide_legend(title.position = "top",title.hjust = 0.5))+
  #annotate("label", x = 0.7, y = 0, vjust = 1.3, label = "/ /",fill="white")+
theme_light()+
  theme(  axis.title = element_text(colour = "grey30"),
          axis.text = element_text(colour = "grey50"),
           #axis.ticks.x = element_line(colour = "grey90"),
          panel.border = element_blank(),
          axis.line = element_line(colour = "grey80"),
           axis.text.y = element_blank(),
          axis.line.y = element_blank(),
          panel.grid.major = element_line(colour = "grey90"),
          legend.text = element_text(colour = "grey50"),
          legend.title = element_text(colour = "grey30"),
          legend.position="top")
p_G_sp_mortality_agric
```

### 1.2.3 Plot of abundances of Gammarus fossarum
```{r}
str(df2)
df2$site_type=tolower(df2$site_type)
df_abundance<-df2 %>% dplyr::select(c("stream","site_type","Gammarus_fossarum","Gammarus_pulex","Gammarus_roeselii"))

#plot abundance
fig_abundance <- 
    ggplot(df_abundance, aes(x = factor(site_type,level=c("refuge","edge","agriculture")) , y = Gammarus_fossarum,fill=site_type)) + 
        geom_point(mapping=aes(colour=site_type,alpha=site_type),size=3,position=position_jitterdodge())+#) 
  geom_hline(yintercept=50, linetype="dashed", color = "steelblue")+
  geom_hline(yintercept=80, linetype="dashed", color = "red",alpha=0.6)+
   annotate(geom="text", x=0.85, y=83, label="↑ High", color="red",alpha=0.6)+
 annotate(geom="text", x=0.85, y=48, label="↓ Low", color="steelblue")+
         labs(title="",x="Site type", y = expression(paste("Relative abundance of" , italic(" Gammarus fossarum")))) +
       scale_colour_manual(values=group.color,name="Site type",breaks = c("refuge","edge","agriculture"), labels=c("Refuge","Edge","Agriculture"))+
    scale_fill_manual(values=group.color,name="Site type",breaks = c("refuge","edge","agriculture"), labels=c("Refuge","Edge","Agriculture"))+
  scale_alpha_manual(values=c(0.4,0.75,0.6),name="Site type",breaks = c("refuge","edge","agriculture"), labels=c("Refuge","Edge","Agriculture"))+
     scale_x_discrete(name="Site type",breaks = c("refuge","edge","agriculture"), labels=c("Refuge","Edge","Agriculture"))+
  scale_y_continuous(breaks = c(0,20,40,60,80,100), labels=c("0%","20%","40%","60%","80%","100%"),limits = c(-1,102))+
     theme_light()+
  theme(axis.title = element_text(colour = "grey30"),
          axis.text = element_text(colour = "grey50"),
           axis.text.x = element_text(angle = 10, vjust = 0.8),
          panel.border = element_blank(),
          axis.line = element_line(colour = "grey80"),
          panel.grid.major = element_line(colour = "grey90"),
          legend.text = element_text(colour = "grey50"),
          legend.title = element_text(colour = "grey30"),
          legend.position="none")
fig_abundance
```


### 1.2.4 Immobility after 96h imidacloprid exposure - Predictor Site type
#### 1.2.4.1 Statistics
a) Modelling
```{r echo=FALSE}
#statistics for the following dataframe
df_statistics<-df_gam %>% dplyr::select(c("Dead_96","Individuals","LogConcentration","fSiteType","fStream","fGFossarumAbundance","Immobile_96","Alive_96","fSubunit"))
str(df_statistics)
```

```{r}
#note, per definition immobile means in our study, still alive, thus we add the dead individuals for an overall response
#calculate response (dead+immobile)
df_statistics$response_96<-df_statistics$Dead_96+df_statistics$Immobile_96 
#calculate proportion of response
df_statistics$Proportion_response_96<-df_statistics$response_96/df_statistics$Individuals # calculate response proportion

# remove NAs
df_statistics<-df_statistics[!is.na(df_statistics$response_96),]
unique(df_statistics$response_96)
```

a) Data exploration
```{r}
#outlier for response and continuous covariates
MyVar <- c("Proportion_response_96","LogConcentration")
Mydotplot(df_statistics[,MyVar])
outlier_values <- boxplot.stats(df_statistics$Proportion_response_96)$out  # outlier values.
boxplot(df_statistics$Proportion_response_96, main="outliers", boxwex=0.1)
mtext(paste("Outliers: ", paste(outlier_values, collapse=", ")), cex=0.6)
#variance homogeneity
p_vh<-beanplot(Proportion_response_96~fSiteType, data=df_statistics)
leveneTest(Proportion_response_96~fSiteType, data=df_statistics)
#normal distribution
hist(df_statistics$Proportion_response_96)
shapiro.test(df_statistics$Proportion_response_96)
#Inspect for collinearity with multi-panel scatterplots (only covariates): 
MyVar1 <- c("LogConcentration","fGFossarumAbundance","fSiteType") # multi-panel scatterplots
pairs(df_statistics[,MyVar1], 
    upper.panel = panel.cor,
  diag.panel  = panel.hist,
  lower.panel = panel.smooth)
#Inspect for collinearity with VIF, cut-off level in Zuur et al., 2013 VIF = 3
corvif(df_statistics[ ,c("LogConcentration","fGFossarumAbundance","fSiteType")]) # calculate variance inflation factors(VIFs) for fixed covariates
```
Data are not normally distributed,Binary data. Variances homogenize and there are no problematic outliers.

b) Modeling
```{r echo=FALSE}
#Executing the GLMM using the glmer function
Full_model <- glmer(Proportion_response_96 ~ LogConcentration*fSiteType+(1|fStream)+(1|fSubunit),weights = Individuals,
            family = binomial, data = df_statistics,
            glmerControl(optimizer = "optimx", calc.derivs = FALSE,
                         optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)))
```

b) Model selection
```{r}
Anova(Full_model)
drop1(Full_model)#Interaction is not significant, we remove it from the model.

Reduced_model <- glmer(Proportion_response_96 ~ LogConcentration+fSiteType+(1|fStream)+(1|fSubunit),weights = Individuals,
            family = binomial, data = df_statistics,
            glmerControl(optimizer = "optimx", calc.derivs = FALSE,
                         optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)))
Anova(Reduced_model)
drop1(Reduced_model)#fSiteType is not significant
```

c) Model validation
```{r}
#model validation with the performance package
check_model(Reduced_model) # ok
model_performance(Reduced_model) 
check_overdispersion(Reduced_model) 
check_singularity(Reduced_model) #singular - random factor stream variances not available, shrinked to zero
check_collinearity(Reduced_model) 
```

e) Model output
```{r}
summary(Reduced_model)
# create a nice output
tab_model(Reduced_model,show.ci = F, show.se = TRUE)#,  string.se = "SE")

#significance of fixed effects
Anova(Reduced_model,type=2)#concentration is significant
drop1(Reduced_model)
#store in unique name
Model3<-Reduced_model
```
Result: No significant difference between site types.

#### 1.2.4.2 EC50
```{r}
#calculate EC50
function_calculate_LC50 = function(b0,b1,b2,b3,prob){
x1 = (-b0-b2-b3+log(-prob/(prob-1)))/b1
#x1 = (log(prob/(1-prob))-b0-b2-b3)/b1 #both lines are the same
return(10^(x1)) # inverse of log10() for the concentration, given that it was log10(x+1) transformed
}

#access the estimated parameters and calculate predictor function
ff<-fixef(Reduced_model)
LC_all<-function_calculate_LC50(ff[[1]],ff[[2]],ff[[3]],ff[[4]],.5)
LC_agriculture<-function_calculate_LC50(ff[[1]],ff[[2]],0,0,.5) #LC50 agriculture
LC_edge<-function_calculate_LC50(ff[[1]],ff[[2]],ff[[3]],0,.5) #LC50 edge
LC_refuge<-function_calculate_LC50(ff[[1]],ff[[2]],0,ff[[4]],.5) #LC50 refuge
LC_agriculture # 72 µg/L
LC_edge # 87 µg/L
LC_refuge # 77 µg/L
```

#### 1.2.1.3 Plot
```{r}
#Predict new data
#create grid of covariate values and get dummy variables
NewData <- expand.grid(fSiteType=c("agriculture","edge","refuge"),
                        LogConcentration=seq(log10(2.3),3.4,length=1000))
#figure out the model matrix X corresponding to the new data
X <- model.matrix(~ LogConcentration + fSiteType, data = NewData)

#access the estimated parameters and calculate predictor function
ff<-fixef(Reduced_model)
NewData$eta         <- X %*% ff

#calculate the probability of mortality for the specified covariate values using the inverse of the logistic link function
NewData$SuccessPred <- exp(NewData$eta) / (1 + exp(NewData$eta)) 

#extract the variance-covariance matrix of the parameter V , compute XVX' to get the variance-covariance matrix of the predictions, then extract the diagonal of this matrix to get variances of predictions
output<-vcov(Reduced_model) #use only for glmmTMB models,called "variance"
NewData$VarPred <- diag(X %*% output %*% t(X))#unconditional variance
#take the sqare-root to get standard deviations (errors) of the predictions
NewData$sePred  <- sqrt(NewData$VarPred)

#obtain 95% confidence intervals on the scale of the response variable
#for GLMMs, run the confidence interval boundaries (not the standard errors) through the inverse-link function
NewData$seLow  <- exp(NewData$eta - 1.96 * NewData$sePred) / (1 + exp(NewData$eta - 1.96 * NewData$sePred))
NewData$seHigh <- exp(NewData$eta + 1.96 * NewData$sePred) / (1 + exp(NewData$eta + 1.96 * NewData$sePred))

#aggregate data as mean of all subunit replicates
df_mean<-aggregate(list(df_statistics$Proportion_response_96, df_statistics$Individuals), by=list(df_statistics$fSubunit, df_statistics$fSiteType,df_statistics$fStream,df_statistics$LogConcentration,df_statistics$fGFossarumAbundance), mean)
str(df_mean)
names(df_mean)<-c("fSubunit","fSiteType","fStream","LogConcentration","fGFossarumAbundance","Proportion_response_96","Individuals")

#use this one if you want to show the mean of all subunit replicates, to avoid overplotting
df_statistics2  <-df_mean %>% mutate(result=NA, result=ifelse(fSiteType =="agriculture",LogConcentration-0.04,result), result=ifelse(fSiteType =="refuge",LogConcentration+0.04,result), result=ifelse(fSiteType =="edge",LogConcentration,result)  )

p_G_sp_immobility <- 
  ggplot(NewData, mapping=aes(x=LogConcentration, y=SuccessPred,group=fSiteType)) +
  geom_point(df_statistics2,mapping = aes(x=result,y=Proportion_response_96,colour=fSiteType, alpha = fSiteType),size=1.5)+ # position = position_dodge2(preserve = "single")  
  geom_line(aes(colour=fSiteType),size=1.2, alpha=1) +
    geom_ribbon(aes( ymin= seLow, ymax=seHigh, fill = fSiteType, alpha = fSiteType) )+
     scale_x_continuous(breaks = c(log10(1),log10(10),log10(100),log10(1000),log10(5000)), labels=c(1,10,100,1000,5000),limits = c(log10(1),log10(5000)))+
  scale_y_continuous(breaks = c(0,0.20,0.40,0.6,0.80,1), labels=c("0%","20%","40%","60%","80%","100%"),limits = c(0,1))+
#  scale_colour_manual()+
  scale_colour_manual(values=group.color,name="Site type",breaks = c("refuge","edge","agriculture"), labels=c("Refuge","Edge","Agriculture"))+
    scale_fill_manual(values=group.color,name="Site type",breaks = c("refuge","edge","agriculture"), labels=c("Refuge","Edge","Agriculture"))+ 
   scale_alpha_manual(values=c(0.2,0.4,0.3),name="Site type",breaks = c("refuge","edge","agriculture"), labels=c("Refuge","Edge","Agriculture"))+
  annotation_logticks(sides = "b",colour = "grey50")+
    labs(x = expression(paste("Imidacloprid concentration [µg/L]")), y = 'Immobility')+
    theme_light()+
  theme(  axis.title = element_text(colour = "grey30"),
          axis.text = element_text(colour = "grey50"),
           #axis.ticks.x = element_line(colour = "grey90"),
          panel.border = element_blank(),
          axis.line = element_line(colour = "grey80"),
          panel.grid.major = element_line(colour = "grey90"),
          legend.text = element_text(colour = "grey50"),
          legend.title = element_text(colour = "grey30"),
          legend.position="right")
p_G_sp_immobility
```

## 1.3 Plot Tolerance results all together
```{r}
#save Figure 
dev.new()
figure_1<-ggarrange(nrow=1,p_G_sp_mortality,p_G_sp_mortality_agric+rremove("ylab"),fig_abundance,widths = c(1,0.85,0.6),  labels = c("A","B","C") )#, )#)align = "h",
# ggsave(path = Outpath,filename = "20221121_Figure_1.eps",figure_1, width=9.5, height=4.5,pointsize=20,device=cairo_ps, fallback_resolution = 600)
ragg::agg_tiff(paste(Outpath,"20230511_Figure_1.tiff",sep="/"), width = 160, height = 70, units = "mm", res = 300, scaling = 0.7)
figure_1
dev.off()

dev.new()
#ggsave(path = Outpath,filename = "20221121_Figure_SI_2.eps",p_G_sp_immobility, width=6, height=4, pointsize=15,device=cairo_ps, fallback_resolution = 600)
ragg::agg_tiff(paste(Outpath,"20230505_Figure_SI_2.tiff",sep="/"), width = 5, height = 3.5, units = "in", res = 600, scaling = 0.8)
p_G_sp_immobility
dev.off()
```


```{r}
tab_model(Model1,Model2,Model3,show.ci = F, show.se = TRUE)
```


