---
title: "Population structure and insecticide response of Gammarus spp. in agricultural and upstream forested sites of small streams "
author: "Schneeweiss A., Schreiner V.C., Liess M., Roeder N., Schwenk K., Schaefer R.B."
date: "01 November 2022"
output:
  html_document:
    figure_caption: yes
    highlight: tango
    number_sections: no 
    toc: yes
    toc_float: 
         toc_collapsed: true
    toc_depth: 4
  pdf_document:
    toc: yes
    toc_float: 
         toc_collapsed: true
    toc_depth: '4'
  word_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

R script written by Anke Schneeweiss

Internally verified by Dr. Verena C. Schreiner

Institute for Environmental Sciences
RPTU Kaiserslautern-Landau 
Fortstrasse 7  
76829 Landau  
GERMANY  
Corresponding mail address: schneeweiss'@'uni-landau.de

```{r, include=FALSE}
#Required packages
require(plyr) 
require(dplyr) 
require(tidyr) 
require(stringr)
require(data.table) 
require(reshape) 
require(ggplot2) 
require(vegan) 
require(EnvStats)
require(ggpubr)
require(scales)
require(gridExtra)
require(ggpmisc)
require(reactable) 
require(kableExtra) 
require(qwraps2)
require(doBy) 
require(knitr)
require(tidyselect)
require(janitor)
require(lme4)
require(beanplot)
require(lmerTest)
require(car)
require(performance)
require(sjPlot)
require(DHARMa)
require(multcomp)
require(emmeans)
require(DAAG) 
require(effects) 
require(pbkrtest)
library(ggeffects)
#source(file = ".../HighstatLibV10.R") #download file "GLMGLMM_RCode.zip" on https://www.highstat.com/index.php/books2?view=article&id=21&catid=18 and read in "HighstatLibV10.R" from local source TODO
library('RVAideMemoire')
```


```{r, include=FALSE}
#########Set the working directory ##########
path<-getwd()
#Set data input path
Inpath<-paste(path,"0_data_input" ,sep="/")
#set analysis output path
Outpath<-paste(path,"1_analysis_output" ,sep="/")
```


# 0. Some general definitions
We define a fixed color scheme for the site types.
```{r}
#group color scheme for site_types in all plots
group.color <- c("agriculture"= "peru", "edge" ="darkkhaki" , "refuge" = "darkgreen")
```
Rounds 0.5 up to 1.
```{r}
round2 <- function(x, digits = 0) {  
  posneg <- sign(x)
  z <- abs(x) * 10^digits
  z <- z + 0.5
  z <- trunc(z)
  z <- z / 10^digits
  z * posneg
}
```


# 1. Hypothesis 3: Estimating energy reserves
## 1.1 Load and prepare data
```{r}
df4 <- read.csv(file=paste(Inpath,"20230505_data_part_3_lipids.csv", sep="/"),header=TRUE, sep=",", na.strings =c("na","NA",""),stringsAsFactors = FALSE)
df5<-read.csv(file=paste(Inpath,"20230505_data_4_gammarus_abundance.csv", sep="/"),header=TRUE, sep=";", na.strings ="na",stringsAsFactors = FALSE)

str(df4)
str(df5)
##renaming
df5$site_type=tolower(df5$site_type)
df5 <- unite(df5,"Sampling_site",c("stream","site_type"),sep="_",remove = F)
```
Data set:
df1: data of lipid analysis from two separate runs for 18 sites. 
Run 1: All Species (G.fossarum, G.pulex and G.roeseli) can occur. 
Run 2: Only G.fossarum. 0-5 organism replicates per sampling site, according to the availability of G.fossarum. At some sampling sites (e.g. Krops_down, Russ_down) G.fossarum does not occur. Therefore, we have less replicates for the agricultural site type.
df2: data for the relative abundance of Gammarus fossarum at the 18 sampling site (high abundance: >=80% Gammarus fossarum; low: <=50% Gammarus fossarum)

a) Subset two individual dataframes, one for run 1 and one for run 2
```{r}
df_energy_reserves<-df4
df_run1<-subset(df_energy_reserves,analysis_date=="20.01.2020"|analysis_date=="22.01.2020"|analysis_date=="24.01.2020"|analysis_date=="27.01.2020"|analysis_date=="29.01.2020")
df_run1$run<-"use_for_1"
df_run1$identified_species<-"Gammarus sp"

df_run2<-subset(df_energy_reserves,analysis_date=="02.03.2020"|analysis_date=="04.03.2020"|analysis_date=="06.03.2020"|analysis_date=="01.07.2020")
df_run2 <- subset(df_run2,identified_species=="Gammarus fossarum")# Select only Gammarus fossarum individuals - species identification according to genetic CO1 results
df_run2$run<-"use_for_2"
str(df_run1)
str(df_run2)
#df_complete<-rbind(df_run1,df_run2)
```

## 1.2 Run 1
### 1.2.1 Weight
To avoid growth effects, we normalised the lipid content to the gammarids' dry weight (µg/mg gammarid) (Zubrod et al., 2011). Therefore, we measured and analysed the individuals dry weight first: five individuals  per sampling site were freeze-dried for 24 h and subsequently weighed to the nearest 0.001 mg.
#### 1.2.1.1 Statistics ~ Gammarus abundance
```{r}
#statistics for the following dataframe
df_statistics<-df_run1
str(df_statistics)

#add info "abundance" of Gammarus abundance
df_statistics <- unite(df_statistics,"Sampling_site",c("stream","site_type"),sep="_",remove = F)
df_statistics<-join(df_statistics,df5[,c("Sampling_site","Gammarus_fossarum","abundance")],by = "Sampling_site")

#change variables to factor
df_statistics$site_type  	<- factor(df_statistics$site_type) 
df_statistics$stream  	<- factor(df_statistics$stream)
df_statistics <-  unite(df_statistics,"replicate",c("stream","site_type"),sep="_",remove = F)
df_statistics$replicate  	<- factor(df_statistics$replicate)
df_statistics$fGFossarumAbundance	<- factor(df_statistics$abundance) #"abundance" describes the relative abundance of Gammarus fossarum at the respective sampling site (high abundance: >=80% Gammarus fossarum; low: <=50% Gammarus fossarum)
```

a) Data exploration
```{r}
#Observe number of observation - Balanced design
df_statistics %>% count(replicate,sort = TRUE)
str(df_statistics)
#outlier for response and continuous covariates
MyVar <- c("weight_per_individual_mg")
MyVar2 <- c("weight_per_individual_mg","site_type","fGFossarumAbundance")
Mydotplot(df_statistics[,MyVar])
#outlier
boxplot(df_statistics$weight_per_individual_mg, main="outliers", boxwex=0.1)
# boxplot(weight_per_individual_mg~fGFossarumAbundance, data=df_statistics, main="outliers", boxwex=0.1)
#variances
p_vh<-beanplot(weight_per_individual_mg~site_type, data=df_statistics)
leveneTest(weight_per_individual_mg~site_type, data=df_statistics)
p_vh<-beanplot(weight_per_individual_mg~fGFossarumAbundance, data=df_statistics)
leveneTest(weight_per_individual_mg~fGFossarumAbundance, data=df_statistics)
#distribution
hist(df_statistics$weight_per_individual_mg)
byf.hist(weight_per_individual_mg~site_type, data=df_statistics)
byf.hist(weight_per_individual_mg~fGFossarumAbundance, data=df_statistics)
shapiro.test(df_statistics$weight_per_individual_mg)
#Inspect for collinearity with multi-panel scatterplots (only covariates): 
pairs(df_statistics[,MyVar2], lower.panel = panel.cor)
#Inspect for collinearity with VIF, cut-off level in Zuur et al., 2013 VIF = 3
corvif(df_statistics[ ,MyVar2]) # calculate variance inflation factors(VIFs) for fixed covariates
```
Data are not normally distributed, and variances not homogen and there is one outlier. --> However, the model validation based on plots indicates only minor deviation, acceptable

b) Model
```{r}
lmm_run1 <- lmer(weight_per_individual_mg~ fGFossarumAbundance +(1|stream)+(1|replicate), data = df_statistics,REML = T)
Anova(lmm_run1)
drop1(lmm_run1)
#from help('isSingular'): "Complex mixed-effect models (i.e., those with a large number of variance-covariance parameters) frequently result in singular fits, i.e. estimated variance-covariance matrices with less than full rank. Less technically, this means that some "dimensions" of the variance-covariance matrix have been estimated as exactly zero." 
#--> here the random factor stream. We decided to keep it anyway, because it is part of the study design and removing it has negligible influence on parameter estimates.
```


c) Model diagnosis
```{r}
lmm<-lmm_run1
#resid vs fitted values (homogeneity of variances & linear relationship)
plot(lmm)
check_model(lmm)
check_homogeneity(lmm)
#qqplot (normal distribution of error, standardized residuals)
qqnorm(resid(lmm, type = "pearson"))
qqline(resid(lmm, type = "pearson"))
check_normality(lmm) 
#Cook's distance values
plot(cooks.distance(lmm),ylab="Cook distance values")
#general model check
check_singularity(lmm)
check_convergence(lmm) 
```
Model validation based on plots indicates only minor deviation, acceptable.

d) Model output
```{r}
summary(lmm)
tab_model(lmm,show.ci = F, show.se = TRUE, p.val = "kr")
model_performance(lmm) # random factor shrinked to zero
#significance of effects
#F statistics to test fixed effects, Kenward-Roger Approximation is used, it performs well in the range of small sample settings ( see: Kenward MG, Roger JH. Small sample inference for fixed effects from restricted maximum likelihood. Biometrics. 1997 Sep;53(3):983-97. PMID: 9333350.)
anova(lmm,ddf="Kenward-Roger")# no predictor can be droped
emmeans(lmm, list(pairwise ~ fGFossarumAbundance), adjust = "tukey")
model1<-lmm
```

e) Plot
```{r}
fit<-lmm
mydf<-ggeffect(fit,c("fGFossarumAbundance"))

p_weight_gf <-ggplot(mydf,mapping=aes(x, predicted))+
    
  geom_point(df_statistics,mapping=aes(x=fGFossarumAbundance,y=weight_per_individual_mg,colour = fGFossarumAbundance),alpha=0.3,size=2,position=position_jitter(h=0.0,w=0.1))+
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high))+
  geom_signif(comparisons=list(c("low", "high")), annotations="**",
              y_position = 17, tip_length = 0, vjust=0.4) +
  #) 
      labs(title="",x=expression(paste("Abundance of" ,phantom(x), italic("G. fossarum"))), y ="Dry weight per gammarid [mg]") +
   scale_x_discrete(name=expression(paste("Abundance of" ,phantom(x), italic("G. fossarum"))),breaks = c("low","high"), labels=c("Low (≤50%)", "High (≥80%)"))+
  scale_fill_brewer(type = "seq", palette="Set1",name=expression(paste("Abundance of" ,phantom(x), italic("G. fossarum"))),breaks = c("low","high"), labels=c("Low (≤50%, n=4)", "High (≥80%, n=14)"))+
  scale_colour_brewer(type = "seq", palette="Set1",name=expression(paste("Abundance of" ,phantom(x), italic("G. fossarum"))),breaks = c("low","high"), labels=c("Low (≤50%, n=4)", "High (≥80%, n=14)"))+
  scale_y_continuous(breaks = c(0,5,10,15,20), labels=c(0,5,10,15,20),limits = c(0,20))+
  theme_light()+
  theme(  axis.title = element_text(colour = "grey30"),
          axis.text = element_text(colour = "grey50"),
          panel.border = element_blank(),
          axis.line = element_line(colour = "grey80"),
          axis.text.y = element_blank(),
          axis.line.y = element_blank(),
          panel.grid.major = element_line(colour = "grey90"),
          legend.text = element_text(colour = "grey50"),
          legend.title = element_text(colour = "grey30"),
          legend.position="none")
p_weight_gf
```

#### 1.2.1.2 Statistics ~ Site type 
b) Model
```{r}
lmm_run1 <- lmer(weight_per_individual_mg  ~ site_type +(1|stream)+(1|replicate), data = df_statistics,REML = T)
Anova(lmm_run1)
drop1(lmm_run1)
#from help('isSingular'): "Complex mixed-effect models (i.e., those with a large number of variance-covariance parameters) frequently result in singular fits, i.e. estimated variance-covariance matrices with less than full rank. Less technically, this means that some "dimensions" of the variance-covariance matrix have been estimated as exactly zero." 
#--> here the random factor stream. We decided to keep it anyway, because it is part of the study design and removing it has negligible influence on parameter estimates.
```


c) Model diagnosis
```{r}
lmm<-lmm_run1
#resid vs fitted values (homogeneity of variances & linear relationship)
plot(lmm)
check_model(lmm)
check_homogeneity(lmm)
#qqplot (normal distribution of error, standardized residuals)
qqnorm(resid(lmm, type = "pearson"))
qqline(resid(lmm, type = "pearson"))
check_normality(lmm) 
#Cook's distance values
plot(cooks.distance(lmm),ylab="Cook distance values")
#general model check
check_singularity(lmm) 
check_convergence(lmm) 
```
Model validation based on plots indicates only minor deviation, acceptable.

d) Model output
```{r}
summary(lmm)
tab_model(lmm,show.ci = F, show.se = TRUE, p.val = "kr")
model_performance(lmm) #random factors shrinked to zero
#significance of effects
#F statistics to test fixed effects, Kenward-Roger Approximation is used, it performs well in the range of small sample settings ( see: Kenward MG, Roger JH. Small sample inference for fixed effects from restricted maximum likelihood. Biometrics. 1997 Sep;53(3):983-97. PMID: 9333350.)
anova(lmm,ddf="Kenward-Roger")#no predictor can be droped
emmeans(lmm, list(pairwise ~ site_type), adjust = "tukey")
model2<-lmm
```

e) Plot
```{r}
fit<-lmm_run1
mydf<-ggeffect(fit, c("site_type"))

p_weight_st<-ggplot(mydf,mapping=aes(x = factor(x,level=c("refuge","edge","agriculture")), predicted))+#, colour = x
geom_point(df_statistics,mapping=aes(x=factor(site_type,level=c("refuge","edge","agriculture")),y=weight_per_individual_mg,colour =site_type,alpha=site_type),size=2,position=position_jitter(h=0.0,w=0.1))+
    geom_pointrange(aes(ymin = conf.low, ymax = conf.high))+
  geom_signif(comparisons=list(c("refuge", "agriculture")), annotations="*",
              y_position = 17, tip_length = 0, vjust=0.4) +
     geom_signif(comparisons=list(c("edge", "agriculture")), annotations="*",
              y_position = 18.4, tip_length = 0, vjust=0.4) +
      labs(title="",x="Site type", y ="Dry weight per gammarid [mg]") +
 scale_colour_manual(values=group.color,name="Site type",breaks = c("refuge","edge","agriculture"), labels=c("Refuge","Edge","Agriculture"))+
    scale_fill_manual(values=group.color,name="Site type",breaks = c("refuge","edge","agriculture"), labels=c("Refuge","Edge","Agriculture"))+
   scale_alpha_manual(values=c(0.4,0.7,0.5),name="Site type",breaks = c("refuge","edge","agriculture"), labels=c("Refuge","Edge","Agriculture"))+
     scale_x_discrete(name="Site type",breaks = c("refuge","edge","agriculture"), labels=c("Refuge","Edge","Agriculture"))+
   scale_y_continuous(breaks = c(0,5,10,15,20), labels=c(0,5,10,15,20),limits = c(0,20))+
  theme_light()+
  theme(  axis.title = element_text(colour = "grey30"),
          axis.text = element_text(colour = "grey50"),
          panel.border = element_blank(),
          axis.line = element_line(colour = "grey80"),
          panel.grid.major = element_line(colour = "grey90"),
          legend.text = element_text(colour = "grey50"),
          legend.title = element_text(colour = "grey30"),
          legend.position="none")

p_weight_st
```


### 1.2.2 Energy reserves

#### 1.2.2.1 Statistics ~ Gammarus abundance
a) Data exploration
```{r}
#Observe number of observation - Balanced design
df_statistics %>% count(replicate,sort = TRUE)
str(df_statistics)
#outlier for response and continuous covariates
MyVar <- c("lipidcontent_ug_mg")
MyVar2 <- c("lipidcontent_ug_mg","site_type","fGFossarumAbundance")
Mydotplot(df_statistics[,MyVar])
#outlier
#outlier_values <- boxplot.stats(df_statistics$weight_per_individual_mg)$out  # outlier values.
boxplot(lipidcontent_ug_mg~site_type, data=df_statistics, main="outliers", boxwex=0.1)
boxplot(lipidcontent_ug_mg~fGFossarumAbundance, data=df_statistics, main="outliers", boxwex=0.1)
#variances
p_vh<-beanplot(lipidcontent_ug_mg~site_type, data=df_statistics)
leveneTest(lipidcontent_ug_mg~site_type, data=df_statistics)
p_vh<-beanplot(lipidcontent_ug_mg~fGFossarumAbundance, data=df_statistics)
leveneTest(lipidcontent_ug_mg~fGFossarumAbundance, data=df_statistics)
#normal distribution
hist(df_statistics$lipidcontent_ug_mg)
byf.hist(lipidcontent_ug_mg~site_type, data=df_statistics)
byf.hist(lipidcontent_ug_mg~fGFossarumAbundance, data=df_statistics)
shapiro.test(df_statistics$lipidcontent_ug_mg)
#Inspect for collinearity with multi-panel scatterplots (only covariates): 
pairs(df_statistics[,MyVar2], lower.panel = panel.cor)
#Inspect for collinearity with VIF, cut-off level in Zuur et al., 2013 VIF = 3
corvif(df_statistics[ ,MyVar2]) # calculate variance inflation factors(VIFs) for fixed covariates
```


b) Model
```{r}
#linear mixed model
lmm_run1 <- lmer(lipidcontent_ug_mg  ~ fGFossarumAbundance + (1|stream)+ (1|replicate), data = df_statistics,REML = T)
Anova(lmm_run1)
drop1(lmm_run1)
#from help('isSingular'): "Complex mixed-effect models (i.e., those with a large number of variance-covariance parameters) frequently result in singular fits, i.e. estimated variance-covariance matrices with less than full rank. Less technically, this means that some "dimensions" of the variance-covariance matrix have been estimated as exactly zero." 
#--> here the random factor stream. We decided to keep it anyway, because it is part of the study design and removing it has negligible influence on parameter estimates.
```

c) Model diagnosis
```{r}
lmm<-lmm_run1
#resid vs fitted values (homogeneity of variances & linear relationship)
plot(lmm)
check_model(lmm)
check_homogeneity(lmm) 
#qqplot (normal distribution of error, standardized residuals)
qqnorm(resid(lmm, type = "pearson"))
qqline(resid(lmm, type = "pearson"))
check_normality(lmm) 
#Cook's distance values
plot(cooks.distance(lmm),ylab="Cook distance values")
#general model check
check_singularity(lmm) 
check_convergence(lmm)
```
Model validation based on plots indicates only minor deviation, acceptable.

d) Model output TODO
```{r}
summary(lmm)
# create a nice output
tab_model(lmm,show.ci = F, show.se = TRUE, p.val = "kr")
model_performance(lmm) # random factors describes some variance
#significance of effects
#F statistics to test fixed effects, Kenward-Roger Approximation is used, it performs well in the range of small sample settings ( see: Kenward MG, Roger JH. Small sample inference for fixed effects from restricted maximum likelihood. Biometrics. 1997 Sep;53(3):983-97. PMID: 9333350.)
anova(lmm,ddf="Kenward-Roger")# not significant
#posthoc - more conservative --> used for main manuscript
emmeans(lmm, list(pairwise ~ fGFossarumAbundance), adjust = "tukey")
model3<-lmm
```
Result: We did not find evidence that the level of energy reserves differs across G.fossarum abundance (LMM, factor site type not significant at p = 0.115).

e) Plot
```{r}
fit<-lmm
mydf<-ggeffect(fit, c( "fGFossarumAbundance"))

p_energy_gf<-ggplot(mydf,mapping=aes(x = x, predicted))+#, colour = x
geom_point(df_statistics,mapping=aes(x=fGFossarumAbundance,y=lipidcontent_ug_mg,colour =fGFossarumAbundance),alpha=0.3,size=2,position=position_jitter(h=0.0,w=0.1))+
   geom_pointrange(aes(ymin = conf.low, ymax = conf.high))+#) 
      labs(title="",x=expression(paste("Abundance of" ,phantom(x), italic("G. fossarum"))), y ="Lipid content [µg/mg gammarid]") +
  scale_x_discrete(name=expression(paste("Abundance of" ,phantom(x), italic("G. fossarum"))),breaks = c("low","high"), labels=c("Low (≤50%)", "High (≥80%)"))+
  scale_fill_brewer(type = "seq", palette="Set1",name=expression(paste("Abundance of" ,phantom(x), italic("G. fossarum"))),breaks = c("low","high"), labels=c("Low (≤50%, n=4)", "High (≥80%, n=14)"))+
  scale_colour_brewer(type = "seq", palette="Set1",name=expression(paste("Abundance of" ,phantom(x), italic("G. fossarum"))),breaks = c("low","high"), labels=c("Low (≤50%, n=4)", "High (≥80%, n=14)"))+
  scale_y_continuous(breaks = c(0,50,100,150,200,250), labels=c(0,50,100,150,200,250),limits = c(0,250))+
  theme_light()+
  theme(  axis.title = element_text(colour = "grey30"),
          axis.text = element_text(colour = "grey50"),
          panel.border = element_blank(),
          axis.line = element_line(colour = "grey80"),
           axis.text.y = element_blank(),
          axis.line.y = element_blank(),
          panel.grid.major = element_line(colour = "grey90"),
          legend.text = element_text(colour = "grey50"),
          legend.title = element_text(colour = "grey30"),
          legend.position="none")

p_energy_gf
```

#### 1.2.2.2 Statistics ~ Site type
b) Model
```{r}
#linear mixed model site type
lmm_run1 <- lmer(lipidcontent_ug_mg  ~ site_type + (1|stream)+ (1|replicate), data = df_statistics,REML = T)
Anova(lmm_run1)
drop1(lmm_run1)
```

c) Model diagnosis
```{r}
lmm<-lmm_run1
#resid vs fitted values (homogeneity of variances & linear relationship)
plot(lmm)
check_model(lmm)
check_homogeneity(lmm)
#qqplot (normal distribution of error, standardized residuals)
qqnorm(resid(lmm, type = "pearson"))
qqline(resid(lmm, type = "pearson"))
check_normality(lmm) 
#Cook's distance values
plot(cooks.distance(lmm),ylab="Cook distance values")
#general model check
check_singularity(lmm) 
check_convergence(lmm)
```
Model validation based on plots indicates only minor deviation, acceptable.

d) Model output
```{r}
summary(lmm)
tab_model(lmm,show.ci = F, show.se = TRUE,   p.val = "kr")
model_performance(lmm) # random factors describes some variance
#significance of effects
#F statistics to test fixed effects, Kenward-Roger Approximation is used, it performs well in the range of small sample settings ( see: Kenward MG, Roger JH. Small sample inference for fixed effects from restricted maximum likelihood. Biometrics. 1997 Sep;53(3):983-97. PMID: 9333350.)
anova(lmm,ddf="Kenward-Roger")#site_type is not significant
#posthoc - more conservative --> used for main manuscript
emmeans(lmm, list(pairwise ~ site_type), adjust = "tukey")
model4<-lmm
```
Conclusion: We did not find evidence that the level of energy reserves differs across site type (LMM, factor site type not significant at p = 0.29).

e) Plot
```{r}
fit<-lmm
mydf<-ggeffect(fit, c("site_type"))

p_energy_st<-ggplot(mydf,mapping=aes(x = factor(x,level=c("refuge","edge","agriculture")), predicted))+#, colour = x
    
geom_point(df_statistics,mapping=aes(x=factor(site_type,level=c("refuge","edge","agriculture")),y=lipidcontent_ug_mg,colour =site_type,alpha=site_type),size=2,position=position_jitter(h=0.0,w=0.1))+
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high))+#) 
      labs(title="",x="Site type", y ="Lipid content [µg/mg gammarid]") +
 scale_colour_manual(values=group.color,name="Site type",breaks = c("refuge","edge","agriculture"), labels=c("Refuge","Edge","Agriculture"))+
    scale_fill_manual(values=group.color,name="Site type",breaks = c("refuge","edge","agriculture"), labels=c("Refuge","Edge","Agriculture"))+
  scale_alpha_manual(values=c(0.4,0.7,0.5),name="Site type",breaks = c("refuge","edge","agriculture"))+
     scale_x_discrete(name="Site type",breaks = c("refuge","edge","agriculture"), labels=c("Refuge","Edge","Agriculture"))+
  scale_y_continuous(breaks = c(0,50,100,150,200,250), labels=c(0,50,100,150,200,250),limits = c(0,250))+
  theme_light()+
  theme(  axis.title = element_text(colour = "grey30"),
          axis.text = element_text(colour = "grey50"),
          panel.border = element_blank(),
          axis.line = element_line(colour = "grey80"),
          panel.grid.major = element_line(colour = "grey90"),
          legend.text = element_text(colour = "grey50"),
          legend.title = element_text(colour = "grey30"),
          legend.position="none")
p_energy_st
```

## 1.3 Run 2
Run 2: Only G.fossarum. 0-5 organism replicates per sampling site, according to the availability of G.fossarum. At some sampling sites (e.g. Krops_down, Russ_down) G.fossarum does not occur. Therefore, we have less replicates for the agricultural site type.

```{r}
#statistics for the following dataframe
df_statistics<-df_run2
str(df_statistics)
#change variables to factor
df_statistics$site_type  	<- factor(df_statistics$site_type)
df_statistics$stream  	<- factor(df_statistics$stream)
df_statistics <-unite(df_statistics,"replicate",c("stream","site_type"),sep="_",remove = F)
df_statistics$replicate  	<- factor(df_statistics$replicate)
```

### 1.3.1 Weight
a) Data exploration
```{r}
#Observe number of observation - Balanced design
df_statistics %>% count(replicate,sort = TRUE)
str(df_statistics)
#outlier for response and continuous covariates
MyVar <- c("weight_per_individual_mg")
MyVar2 <- c("weight_per_individual_mg","site_type")
Mydotplot(df_statistics[,MyVar])
#outlier
boxplot(weight_per_individual_mg~site_type, data=df_statistics, main="outliers", boxwex=0.1)
#variances
p_vh<-beanplot(weight_per_individual_mg~site_type, data=df_statistics)
leveneTest(weight_per_individual_mg~site_type, data=df_statistics)
#normal distribution
hist(df_statistics$weight_per_individual_mg)
byf.hist(weight_per_individual_mg~site_type, data=df_statistics)
shapiro.test(df_statistics$weight_per_individual_mg)
#Inspect for collinearity with multi-panel scatterplots (only covariates): 
pairs(df_statistics[,MyVar2], lower.panel = panel.cor)
#Inspect for collinearity with VIF, cut-off level in Zuur et al., 2013 VIF = 3
corvif(df_statistics[ ,MyVar2]) 
```
Post model validation based on plots indicates only minor deviation, acceptable

b) Model
```{r}
lmm_run2 <- lmer(weight_per_individual_mg ~ site_type +(1|stream)+(1|replicate), data = df_statistics,REML = T)
Anova(lmm_run2)
drop1(lmm_run2)
```

c) Model diagnosis
```{r}
lmm<-lmm_run2
#resid vs fitted values (homogeneity of variances & linear relationship)
plot(lmm)
check_model(lmm)
#qqplot (normal distribution of error, standardized residuals)
qqnorm(resid(lmm, type = "pearson"))
qqline(resid(lmm, type = "pearson"))
#Cook's distance values
plot(cooks.distance(lmm),ylab="Cook distance values")
#general model check
check_singularity(lmm) 
check_convergence(lmm) 
```
Model is validated.

d) Model output
```{r}
summary(lmm)
# create a nice output
tab_model(lmm,show.ci = F, show.se = TRUE,   p.val = "kr")
model_performance(lmm) # random factors describes variance
#significance of effects
#F statistics to test fixed effects, Kenward-Roger Approximation is used, it performs well in the range of small sample settings ( see: Kenward MG, Roger JH. Small sample inference for fixed effects from restricted maximum likelihood. Biometrics. 1997 Sep;53(3):983-97. PMID: 9333350.)
anova(lmm,ddf="Kenward-Roger")#site_type is not significant
emmeans(lmm, list(pairwise ~ site_type), adjust = "tukey")
model5<-lmm
```

e) Plot
```{r}
fit<-lmm_run2
mydf<-ggeffect(fit, c("site_type"))

p_weight_2<-ggplot(mydf,mapping=aes(x = factor(x,level=c("refuge","edge","agriculture")), predicted))+#, colour = x
geom_point(df_statistics,mapping=aes(x=factor(site_type,level=c("refuge","edge","agriculture")),y=weight_per_individual_mg,colour =site_type,alpha=site_type),size=2,position=position_jitter(h=0.0,w=0.1))+
    geom_pointrange(aes(ymin = conf.low, ymax = conf.high))+ 
      labs(title="",x="Site type", y ="Dry weight per gammarid [mg]") +
 scale_colour_manual(values=group.color,name="Site type",breaks = c("refuge","edge","agriculture"), labels=c("Refuge","Edge","Agriculture"))+
    scale_fill_manual(values=group.color,name="Site type",breaks = c("refuge","edge","agriculture"), labels=c("Refuge","Edge","Agriculture"))+
  scale_alpha_manual(values=c(0.4,0.7,0.5),name="Site type",breaks = c("refuge","edge","agriculture"), labels=c("Refuge","Edge","Agriculture"))+
     scale_x_discrete(name="Site type",breaks = c("refuge","edge","agriculture"), labels=c("Refuge","Edge","Agriculture"))+
  theme_light()+
  theme(  axis.title = element_text(colour = "grey30"),
          axis.text = element_text(colour = "grey50"),
          panel.border = element_blank(),
          axis.line = element_line(colour = "grey80"),
          panel.grid.major = element_line(colour = "grey90"),
          legend.text = element_text(colour = "grey50"),
          legend.title = element_text(colour = "grey30"),
          legend.position="none")
p_weight_2
```

### 1.3.2 Energy reserves
a) Data exploration
```{r}
#Observe number of observation - Balanced design
df_statistics %>% count(replicate,sort = TRUE)
#outlier
boxplot(lipidcontent_ug_mg~site_type, data=df_statistics, main="outliers", boxwex=0.1)
#variances
p_vh<-beanplot(lipidcontent_ug_mg~site_type, data=df_statistics)
leveneTest(lipidcontent_ug_mg~site_type, data=df_statistics)
#normal distribution
hist(df_statistics$weight_per_individual_mg)
byf.hist(lipidcontent_ug_mg~site_type, data=df_statistics)
shapiro.test(df_statistics$lipidcontent_ug_mg)
#outlier for response and continuous covariates
MyVar <- c("lipidcontent_ug_mg")
MyVar2 <- c("lipidcontent_ug_mg","site_type")
Mydotplot(df_statistics[,MyVar])
#Inspect for collinearity with multi-panel scatterplots (only covariates): 
pairs(df_statistics[,MyVar2], lower.panel = panel.cor)
#Inspect for collinearity with VIF, cut-off level in Zuur et al., 2013 VIF = 3
corvif(df_statistics[ ,MyVar2])
```

b) Model
```{r}
#linear mixed model
lmm_run2 <- lmer(lipidcontent_ug_mg ~ site_type + (1|stream) +(1|replicate), data = df_statistics,REML = T)
Anova(lmm_run2)
drop1(lmm_run2)
```

c) Model diagnosis
```{r}
lmm<-lmm_run2
#resid vs fitted values (homogeneity of variances & linear relationship)
plot(lmm)
check_model(lmm)
check_homogeneity(lmm) 
#qqplot (normal distribution of error, standardized residuals)
qqnorm(resid(lmm, type = "pearson"))
qqline(resid(lmm, type = "pearson"))
check_normality(lmm) 
#Cook's distance values
plot(cooks.distance(lmm),ylab="Cook distance values")
#general model check
check_singularity(lmm) 
check_convergence(lmm) 
```
Model is validated.

d) Model output
```{r}
summary(lmm)
tab_model(lmm,show.ci = F, show.se = TRUE,  p.val = "kr")
model_performance(lmm) # random factors describes variance
#significance of effects
#F statistics to test fixed effects, Kenward-Roger Approximation is used, it performs well in the range of small sample settings ( see: Kenward MG, Roger JH. Small sample inference for fixed effects from restricted maximum likelihood. Biometrics. 1997 Sep;53(3):983-97. PMID: 9333350.)
anova(lmm,ddf="Kenward-Roger")#site_type is not significant
#posthoc - more conservative --> used for main manuscript
emmeans(lmm, list(pairwise ~ site_type), adjust = "tukey")

model6<-lmm
```
Random factor indicates that both for weight and energy reserves in the second run variate between streams.

e) Plot
```{r}
fit<-lmm
mydf<-ggeffect(fit, c("site_type"))

p_energy_2<-ggplot(mydf,mapping=aes(x = factor(x,level=c("refuge","edge","agriculture")), predicted))+#, colour = x
geom_point(df_statistics,mapping=aes(x=factor(site_type,level=c("refuge","edge","agriculture")),y=lipidcontent_ug_mg,colour =site_type,alpha=site_type),size=2,position=position_jitter(h=0.0,w=0.1))+
 geom_pointrange(aes(ymin = conf.low, ymax = conf.high), alpha=1)+#,fatten = 4
      labs(title="",x="Site type", y ="Lipid content [µg/mg gammarid]") +
 scale_colour_manual(values=group.color,name="Site type",breaks = c("refuge","edge","agriculture"), labels=c("Refuge","Edge","Agriculture"))+
    scale_fill_manual(values=group.color,name="Site type",breaks = c("refuge","edge","agriculture"), labels=c("Refuge","Edge","Agriculture"))+
  scale_alpha_manual(values=c(0.4,0.7,0.5),name="Site type",breaks = c("refuge","edge","agriculture"), labels=c("Refuge","Edge","Agriculture"))+
     scale_x_discrete(name="Site type",breaks = c("refuge","edge","agriculture"), labels=c("Refuge","Edge","Agriculture"))+
  scale_y_continuous(breaks = c(0,50,100,150,200,250), labels=c(0,50,100,150,200,250),limits = c(0,250))+
  theme_light()+
  theme(  axis.title = element_text(colour = "grey30"),
          axis.text = element_text(colour = "grey50"),
          panel.border = element_blank(),
          axis.line = element_line(colour = "grey80"),
          panel.grid.major = element_line(colour = "grey90"),
          legend.text = element_text(colour = "grey50"),
          legend.title = element_text(colour = "grey30"),
          legend.position="none")

p_energy_2
```

## 1.4 Save Figures as tiff
```{r}
#save Figure 3
dev.new()
figure_energy<-ggarrange(p_energy_st,p_energy_gf+ rremove("ylab"),p_weight_st,p_weight_gf+rremove("ylab"),ncol=2,nrow = 2 ,widths = c(1,0.85),labels = c("A","B","C","D")) #labels = c("Gammarus sp","Gammarus fossarum"),label.x = c(0,-0.25)
# ggsave(path = Outpath,filename = "20221121_Figure_4.svg",figure_energy, width=5, height=7, pointsize=15,dpi = 300)
ragg::agg_tiff(paste(Outpath,"20230511_Figure_5.tiff",sep="/"), width = 85, height = 130, units = "mm", res = 300, scaling = 0.7)# width = 3.5, height = 5.5, units = "in"
figure_energy
dev.off()

dev.new()
figure_energy_SI<-ggarrange(p_energy_2,p_weight_2,ncol=2,nrow = 1, labels= c("A","B"))
# ggsave(path = Outpath,filename = "20221121_Figure_SI_4.svg",figure_energy_SI, width=5, height=4, pointsize=15,dpi = 300)

ragg::agg_tiff(paste(Outpath,"20230505_Figure_SI_4.tiff",sep="/"), width = 4.5, height = 3, units = "in", res = 300, scaling = 0.8)
figure_energy_SI
dev.off()

```

## 1.5 Inferential statistics gathered in one table (Table SI 16)
```{r}
#statistic output
tab_model(model1,model2,model3,model4,model5,model6,show.ci = F, show.se = TRUE,  p.val = "kr")#
```





